TacticToe: Learning to Prove with Tactics

### 2018-03-31 tactictoe_jv.tex
switching section 8 and 9
adding psl,ltac,eisbach references

@@

\begin{itemize}
%The different
%recording steps (parsing, unfolding and replaying) are regrouped into a single
%\sml function.
\item Monte Carlo tree search (MCTS) replaces A* as our
  search algorithm (Section~\ref{sec:proofsearch}). The MCTS algorithm gives
more balanced feedback on
the success of
each proof step by
comparing subtrees of the search tree instead of leaves. The policy and
value function are learned
through supervised learning.
\item Proof guidance required by MCTS is given by prediction algorithms
through supervised learning (Section~\ref{s:prediction}).
These predictors are implemented for three kinds of objects: tactics,
theorems, and lists of goals.
\item We introduce an orthogonalization process that eliminates redundant tactics (Section~\ref{sec:ortho}).
\item We introduce a tactic abstraction mechanism
  (Section~\ref{sec:synthesis}), which enables us to create
more general and flexible tactics by dynamically
predicting tactic arguments.
\item The internal ATP \metis is complemented
  by asynchronous calls to \eprover (Section~\ref{sec:atp}) to help \tactictoe
during proof search.
\item Evaluation of \tactictoe with a 60 seconds timeout achieves a 66\%
success rate on the standard library. Comparisons between
\tactictoe and \eprover on different type of problems are reported in
Section~\ref{s:experiments}.
\item Proof recording at the tactic level is made more precise
  (Section~\ref{sec:recording}). We now support pattern
matching constructions and opening SML modules.
\item Minimization and embellishment of the discovered proof facilitates user
  interaction (Section~\ref{sec:proofdisplay}).
\end{itemize}

@@

In Figure~\ref{fig:percentage}, we regroup problems by the length of their
human proof to measure how well \tactictoe and \eprover cope with increasing
levels
of difficulty. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem on their own.
The performance of \tactictoe is compelling with more than half of the theorems
that required a proof of length six being reproven. It is also consistently
better than \eprover for any proof length.

@@

For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15}
can transform a packed tactical proof into a series of interactive tactic
calls. Its principal application
was so far refactoring the library and teaching common proof techniques to new
ITP users. In our work, the splitting of a proof into a sequence of tactics is
essential for the
tactic recording procedure, used to train our tactic prediction mechanism.

@@

% \todoi{Cezary; add ltac and eisbach or remove PSL if you think it is not worth
% mentionning.}
% A similar effort for stronger automation is performed in \isabelle. Nagashima
% developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the
% user to
% program with regular tactics and combine them with tools such as
% \sledgehammer.
% Selection of tactics from data collected in recorded proofs is in development
% but for now it is up to the user to come up with a good meta-tactic.

>>>>

\def\ltac{\textsf{Ltac}\xspace}

@@

\begin{itemize}
%The different
%recording steps (parsing, unfolding and replaying) are regrouped into a single
%\sml function.
\item Monte Carlo tree search (MCTS) replaces A* as our
  search algorithm (Section~\ref{sec:proofsearch}). The MCTS algorithm gives
more balanced feedback on
the success of
each proof step by
comparing subtrees of the search tree instead of leaves. The policy and
value function are learned
through supervised learning.
\item Proof guidance required by MCTS is given by prediction algorithms
through supervised learning (Section~\ref{s:prediction}).
These predictors are implemented for three kinds of objects: tactics,
theorems, and lists of goals.
\item We introduce an orthogonalization process that eliminates redundant tactics (Section~\ref{sec:ortho}).
\item We introduce a tactic abstraction mechanism
  (Section~\ref{sec:synthesis}), which enables us to create
more general and flexible tactics by dynamically
predicting tactic arguments.
\item The internal ATP \metis is complemented
  by asynchronous calls to \eprover (Section~\ref{sec:atp}) to help \tactictoe
during proof search.
\item Proof recording at the tactic level is made more precise
  (Section~\ref{sec:recording}). We now support pattern
matching constructions and opening SML modules.
\item Evaluation of \tactictoe with a 60 seconds timeout achieves a 66\%
success rate on the standard library. Comparisons between
\tactictoe and \eprover on different type of problems are reported in
Section~\ref{s:experiments}.

\item Minimization and embellishment of the discovered proof facilitates user
  interaction (Section~\ref{sec:proofdisplay}).
\end{itemize}

@@

In Figure~\ref{fig:percentage}, we regroup problems by the length of their
human proof to measure how well \tactictoe and \eprover cope with increasing
levels
of difficulty. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem on their own.
The performance of \tactictoe is compelling with more than half of the theorems
that required a proof of length six being re-proven. It is also consistently
better than \eprover for any proof length.

@@

In \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15}
can transform a packed tactical proof into a series of interactive tactic
calls. Its principal application
was so far refactoring the library and teaching common proof techniques to new
ITP users. In our work, the splitting of a proof into a sequence of tactics is
essential for the
tactic recording procedure, used to train our tactic prediction mechanism.

@@

It is important to consider the strength of the language used for building 
proofs and proof procedures when designing a system such as \tactictoe.
In \isabelle, the language {\textsf{Eisbach}\xspace}~\cite{eisbach} was created 
on top of the standard tactic language by giving access to constructors of the 
tactic language to the end-user. The language 
{\textsf{PSL}\xspace}~\cite{NagashimaK17psl} was also developed. It can combine 
regular tactics and tools such as \sledgehammer.
In \coq, the \ltac~\cite{ltac} meta language was designed to enrich the 
tactic combinators of the \coq language. It adds functionalities such as 
recursors and matching operators for terms and proof contexts.
In \holfour, the strength of \sml as a proof strategy 
language has been demonstrated by the implementation of complex proof 
procedures such as \metis.

%Comparatively, in \holfour, new tactics can be written by end-users in the 
%implementation 
%language \sml. Therefore, the design of a meta language was 
%not necessary as every proof method is a \sml function. 
%Using this language, the authors 
%created a default search strategy ``try-hard'' that includes calls to 
%\sledgehammer, and that outperforms \sledgehammer.
